Taking Dyads Seriously

### 2018-02-09 paper/4_app.tex
some updates to the start of the applications section

@@

We choose five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.  The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as decribed in the following sections, the AME model has specific data requirements.}  Each of these pieces was published in a prominent journal and is well-known in the literature. Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables. In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations. We follow the strategy of the original study in our re-estimation of the original models.

>>>>

%% What is the point:

@@

In order to showcase the insights the AME approach permits scholars to bring into their analysis, we illustrate the application of AME on five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.  These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.  We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the method. Firstly, the results of AME estimation are interpretable alongside GLM results. Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, this rich data is increasingly coming online in both new and existing data. Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential. However, reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression. These GLM models ignore dyadic interdependencies, except as they may reveal themselves in included variables, and then adjusting the standard errors in an attempt to account for the clustering of observations.

In the following section, we present two sets of results for each study, one that follows the original study in our re-estimation of the original models and another that tests the central hypothesis within an AME implementation. We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependences in the data.
